---
title: "README"
author: "HL VAN ZYL 17640296"
date: "2025-01-31"
output: github_document
---

# Introduction

This is the README for my financial econometrics 871 project. The primary idea for this project is to compare volatility across market cap sizes. The approach I have decided upon is to decompose the ALSI into three constituent indices each representing different level of market capitalisation and then to compare these indices across various volatility and other statistical measures to gain insight regarding their relative stability in the South African context.

I hope to incorporate the following in my project:

Minimum: 
-Rolling volatility and return plots
-A table containing volatility metrics (RV, Sharpe, maxdrawn downs, tail risks, etc.)


If I have enough time: 
-Stratify over periods of higher USD/ZAR volatility
-Garch model for conditional volatility 

These should provide sufficient evidence to make some inferences regarding the stability of low cap versus mid and large cap stocks in South Africa. 

# Data wrangling 

Load packages
```{r echo=TRUE, message=FALSE, warning=FALSE}
pacman::p_load(tidyverse, here, magrittr, tbl2xts, fmxdat, zoo, PerformanceAnalytics, gt)
```

The data in its raw format has a returns series and also the weights for the ALSI and capped SWIX. For the purposes of this analysis I will be focussing on capped SWIX. Initial data wrangling just cleans the dataframe a bit and removes tickers with more than 3 characters (these don't appear to be legitimate stocks). Impute weights with zeroes where observations are NA. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
df_raw <- readRDS("data/LCL_Stock_Returns.rds") %>% 
  mutate(Tickers = gsub(" SJ Equity", "", Tickers)) %>% 
  rename(Sector = Superindustryname_Pty) %>% 
  arrange(date) %>% 
  rename("ALSI" = "J203", "SWIX"= "J433") %>% 
    select(-VWAP_TO)

# Find tickers with more than 3 characters 
tickers_longer_than_3 <- df_raw %>%
  filter(nchar(Tickers) > 3) %>%
  distinct(Tickers) %>% 
  pull()
# these don't look legit, remove


# use this one for benchmark returns 
# i'm including this  after having done the rest of the project because I thought i
# it might make more sense to have the original ALSI as a benchmark for comparison
# this turned out to be a bad idea ; the values are very close to Large_caps so not 
# very informative and they obscure the other results in the plots
ALSI_bm <- df_raw %>% 
    filter(!Tickers %in% tickers_longer_than_3) %>% 
    mutate(ALSI = coalesce(ALSI,0)) %>% 
    filter(!is.na(Index_Name)) %>% 
    filter(date > lubridate::ymd(20131231)) %>%
    group_by(Tickers, date) %>% 
          mutate(Return = ifelse(is.na(Return),
                         0,
                         Return)) %>% 
    group_by(date) %>% 
    mutate(reweight = ALSI/sum(ALSI)) %>% 
    ungroup()

    


# use this one for market cap return cals
df_raw2 <- df_raw %>% 
    filter(!Tickers %in% tickers_longer_than_3) %>% 
    mutate(ALSI = coalesce(ALSI,0)) %>% 
    filter(!is.na(Index_Name)) %>% 
    select(-SWIX)




```


This next part is a bit tricky. I want to use safe portfolio returns on each Index_Name (market cap size). I create the `calculate_index_returns` to handle this more elegantly and safely via the Safe_Return.portfolio function


```{r}

# import function

source("code/calculate_index_returns.R")

# use function on different mcap sizes

alsi_small <- calculate_index_returns(data = df_raw2, index_name = "Small_Caps", fund_name = "ALSI") %>% 
    rename("Small_Caps"="PortfolioReturn")

alsi_mid <- calculate_index_returns(data = df_raw2, index_name = "Mid_Caps", fund_name = "ALSI") %>% 
    rename("Mid_Caps"="PortfolioReturn")

alsi_large <- calculate_index_returns(data = df_raw2, index_name = "Large_Caps", fund_name = "ALSI") %>% 
    rename("Large_Caps"="PortfolioReturn")

# SWIX ... maybe later

# get them into 1 df

ALSI_returns <- 
    left_join(alsi_small, alsi_mid, by = "date") %>% 
    left_join(., alsi_large, by = "date") %>% 
    # left_join(., ALSI_BM, by = "date") %>%  I'm removing the ALSI benchmark because it obscures the other lines in the plots
    pivot_longer(-date, names_to = "Market_cap", values_to = "Return") %>% 
    filter(date> lubridate::ymd(20131231))

# not sure if it is due to a mistake in my returns calculation or just the nature of the data but pre-2014 data is unusable now; too many dates have been dropped, there are gaps of over 2 months at a time without observations, thus only data starting 2014 will be used for this analysis. 
```

# Rolling plots

Cumulative

```{r}

cumrets <- ALSI_returns %>%
        group_by(Market_cap) %>% 
        mutate(CumReturn = cumprod(1 + Return)) %>% 
        mutate(CumReturn = CumReturn / first(CumReturn)) %>% 
    ungroup()

f <- cumrets %>%  ggplot() + geom_line(aes(date, CumReturn, color = Market_cap),
    alpha = 0.7, size = 1.25) + labs(title = "",
    subtitle = "", x = "", y = "Cumulative Return",
    caption = "Source: Bloomberg and Morningstar") + theme_fmx(title.size = ggpts(30),
    subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = T) +
    fmx_cols()

cum_plot <- finplot(f, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T,
    y.pct = T, y.pct_acc = 1)
    
    
 #Save the plot as a PDF file
# ggsave(
#   filename = here("plots/cum_plot.pdf"), # Replace with your desired folder and file name
#   plot = cum_plot, # The name of your ggplot object
#   device = "pdf",    # The file format
#   width = 7,         # Adjust the width (in inches)
#   height = 4,        # Adjust the height (in inches)
#   units = "in",      # Units for the dimensions
#   dpi = 300          # Dots per inch for high resolution
# )

cum_plot
```


rolling returns 


```{r}
plotdf <-
ALSI_returns%>%
        mutate(YM = format(date, "%Y%B")) %>%
    arrange(date) %>%
    group_by(Market_cap, YM) %>%
    filter(date == last(date)) %>%
    group_by(Market_cap) %>%
    # Epic sorcery:
mutate(RollRets = RcppRoll::roll_prod(1 + Return, 36, fill = NA,
    align = "right")^(12/36) - 1) %>%
    # Note this cool trick: it removes dates that have no
    # RollRets at all.
group_by(date) %>%
    filter(any(!is.na(RollRets))) %>%
    ungroup()

g <-
plotdf %>%
    ggplot() + geom_line(aes(date, RollRets, color = Market_cap),
    alpha = 0.7, size = 1.25) + labs(title = "",
    subtitle = "", x = "", y = "Rolling 3 year Returns (Ann.)",
    caption = "Source: Bloomberg and Morningstar") + theme_fmx(title.size = ggpts(30),
    subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = T) +
    fmx_cols()

ret_plot <- finplot(g, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T,
    y.pct = T, y.pct_acc = 1)


#Save the plot as a PDF file
# ggsave(
#   filename = here("plots/ret_plot.pdf"), # Replace with your desired folder and file name
#   plot = ret_plot, # The name of your ggplot object
#   device = "pdf",    # The file format
#   width = 7,         # Adjust the width (in inches)
#   height = 4,        # Adjust the height (in inches)
#   units = "in",      # Units for the dimensions
#   dpi = 300          # Dots per inch for high resolution
# )

ret_plot
```

rolling sd



```{r}
plot_sd <- ALSI_returns %>%
    mutate(YM = format(date, "%Y%B")) %>%
    arrange(date) %>%
    group_by(Market_cap, YM) %>%
    filter(date == last(date)) %>%
    group_by(Market_cap) %>%
    # Rolling SD annualized calc now:
mutate(RollSD = RcppRoll::roll_sd(1 + Return, 36, fill = NA, align = "right") *
    sqrt(12)) %>%
    filter(!is.na(RollSD))


h <- plot_sd %>%
    ggplot() + geom_line(aes(date, RollSD, color = Market_cap),
    alpha = 0.7, size = 1.25) + labs(title = "",
    subtitle = "", x = "", y = "Rolling 3 year Returns (Ann.)",
    caption = "Source: Bloomberg and Morningstar") + theme_fmx(title.size = ggpts(30),
    subtitle.size = ggpts(5), caption.size = ggpts(25), CustomCaption = T) +
    fmx_cols()

sd_plot <- finplot(h, x.date.dist = "1 year", x.date.type = "%Y", x.vert = T,
    y.pct = T, y.pct_acc = 1)

#Save the plot as a PDF file
# ggsave(
#   filename = here("plots/sd_plot.pdf"), # Replace with your desired folder and file name
#   plot = sd_plot, # The name of your ggplot object
#   device = "pdf",    # The file format
#   width = 7,         # Adjust the width (in inches)
#   height = 4,        # Adjust the height (in inches)
#   units = "in",      # Units for the dimensions
#   dpi = 300          # Dots per inch for high resolution
# )

sd_plot
```


# Comparison tables


Let's first try the Downside risk package
```{r}

wide_xts <- ALSI_returns %>% 
    pivot_wider(., names_from = Market_cap, values_from = Return) %>% tbl2xts::tbl_xts()

tabdownside <-
  table.DownsideRisk(wide_xts, 
                     ci = 0.95, Rf=0, MAR=0)
# Suppose I am only interested in specific elements of the table, then use:
tabdownside <- tabdownside[c(1,5,7,8:11),]
 

gt_tab <- tabdownside %>% data.frame() %>% tibble::rownames_to_column() %>% 
gt() %>% 
        tab_header(title = glue::glue("Downside Risk estimates")) %>% 
      fmt_percent(
      columns = 2:3,
      decimals = 2
    )
#gtsave(gt_tab, here("tables/gt_tab.png"))

#saveRDS(gt_tab, "bin/gt_tab.rds")

gt_tab

```

That's a pretty good start. Next, I want to recreate the statistics stable we made in the practical.

For that, I'll need a benchmark measure. I'll go with the the 2y bond yield.




get 2y bond returns as benchmark comparison
```{r}
sa_bond_2y <- readRDS("data/BondYields_2Y_SA.rds") %>% 
  select(date, Return = BondYield_2) %>% 
  mutate(Bond_2Y = (Return/lag(Return))-1) %>% 
    select(-Return) %>% 
  filter(date> lubridate::ymd(20131231))


Funds <- ALSI_returns %>% 
    pivot_wider(names_from = "Market_cap", values_from = "Return") %>% 
    left_join(sa_bond_2y, by = "date") %>% 
    pivot_longer(-date, names_to = "Tickers", values_to = "Ret") %>% 
    filter(date<lubridate::ymd(20220331))
    
```


Here I just feed my data into the Moments_Comp function which calculates a battery of financial statistics. 

```{r}

source("code/momentscomp.R")

# I'm just going to do the 5 year look back, I had some trouble also including the 3 year

BM = "Bond_2Y"

Tab_stats <- Moments_Comp(funds = Funds, BM = "Bond_2Y", Yrs_LookBack = 5, NA_Check = 0.8)

Make_perc <- 
    
    c( "Cum Returns", "Returns (Ann.)", "Returns Excess (Ann.)", "SD (Ann.)",
       "Avg DD", "Tracking Error")

Rows_to_Perc <- 
    
    Tab_stats %>% mutate(RN=row_number()) %>% filter(Info %in% Make_perc) %>% pull(RN)

colnams <- colnames(Tab_stats)[-1:-2]

Cols_length <- ncol(Tab_stats)


tab <- 
    
    Tab_stats %>% 
    
    gt::gt(groupname_col = 'Period', caption = 'Fund Moments Comparison') %>% 
    tab_header(title = glue::glue("Index Statistics: Relative to {BM}")) %>% 
    fmt_percent(
        columns = 3:Cols_length,
        rows = Rows_to_Perc,
        decimals = 1
    ) %>%   
    
    sub_missing(
        columns = all_of(colnams),
        missing_text = '-'
    ) %>%
    
    tab_footnote( footnote = 'JSAPY is the local property index, consisting of REITS companies.',
                  locations = cells_column_labels(columns = contains('JSAPYTR'))) %>%
    
    tab_footnote(footnote = 'Beta of fund in bull-market periods',
                 locations = cells_column_labels(columns = contains('JSAPYTR'))) %>%
    
    tab_style(
        style = list(
            cell_fill(color = 'gray27', alpha = 0.15),
            cell_text(size = 'large', weight = 'bold',align = 'left')
            
        ),
        locations = cells_row_groups()) 

gt_comp <- tab  %>% 
    
    tab_options(data_row.padding = px(4),table.width = pct(100),
                column_labels.font.size = pct(50),
                column_labels.vlines.width = 1, table.font.size = pct(80)) %>%
    tab_options(data_row.padding = px(6),
                column_labels.font.size = pct(100)) %>% 
    tab_style(style = cell_text(weight = 1200, align = 'left'),locations = cells_title(groups = 'title')) %>%
    tab_style(style = cell_text(color = 'darkgrey', transform = 'uppercase', align = 'center'),
              locations = cells_column_labels(everything()))

# gt::gtsave(tab, here("tables/moments.png"))

#tab 

gt_comp

# gt::gtsave(gt_comp, here("tables/moments.png"))
```

```{r}
library(PerformanceAnalytics)
tablestats <-
  Funds %>% pivot_wider(names_from = "Tickers", values_from = "Ret") %>%  tbl_xts() %>% 
  table.Stats(., ci = 0.95, digits = 3)
#print(tablestats[,1:5])

tablestats
```

Ran out of time, eish. 


